Loop-Aware Optimizations in PyPy's Tracing JIT

### 2011-06-20 talk/iwtc11/paper.tex
kill the limitations section and mention them the conclusion, rewrite the conclusion to be more compact.

\section{Limitations}

XXX as of now?

Loop invariant code motion as described has certain amount of limitations
that prevent it from speeding up larger loops. Those limitations are a target
of future work and might be lifted. Most important ones:

\begin{itemize}
\item Bridges are not well supported - if the flow is more complex than a single
      loop, the bridge might need to jump to the beginning of the preamble,
      making the optimization ineffective
\item XXX write about flushing caches at calls?
\end{itemize}

@@

\subsection{Conclusions}
In this paper we have studied loop invariant code motion during trace
compilation. We claim that loop peeling is a very convenient solution
here since it fits well with other trace optimizations. By peeling of
the first iteration and optimizing the resulting two iteration trace
as a single trace, several standard optimizations can be 
used unchanged. The only interaction needed between the loop peeling
and the other 
optimizations is during the constructing of the jump arguments
connecting the peeled of iteration (the preamble) with the peeled loop. This
approach 
improves the effect of standard optimizations such as redundant guard removal, heap
caching, common subexpression elimination and allocation removals. The
most prominent effect is that they all become loop
invariant code motion optimizations.

By using several benchmarks we show that the proposed algorithm can
significantly
improve the run time of small loops containing numerical
calculations. 
At least in cases where there are not too many guard
failures. A common way of handling a guard that fails often is to 
trace a bridge from it back to the start of some previously compiled
loop. This is applicable here too. However the bridge will have to end
with a jump to the preamble, which lessens the impact of the 
proposed approach.
In many situations it is however possible to make the bridge
jump to the peeled loop instead. When and how this is possible will be
focus of future work.

>>>>

@@

\section{Related Work}
\label{sec:related}

XXX
% section Related Work (end)

\section{Conclusions}

In this paper we have studied loop invariant code motion during trace
compilation. We claim that loop peeling is a very convenient solution
here since it fits well with other trace optimizations and does not require
large changes to them. This approach improves the effect of standard
optimizations such as redundant guard removal, common subexpression elimination
and allocation removal. The most prominent effect is that they all become loop
invariant code motion optimizations.

By using several benchmarks we show that the proposed algorithm can
significantly improve the run time of small loops containing numerical
calculations.

The current approach still has some limitations which we plan to lift in the
future. In particular loop peeling is working less well in combination with
trace trees or trace stitching. The side exits attached guards that fail often
currently have to jump to the preamble which makes loops with several equally
common paths less efficient than they could be.
